---
title: "Personal Experience Data Processing"
---

```{r}
library(readr)
library(zoo)
library(tidyr)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
```


```{r}
data <- read_csv("personalExperienceData.csv",
      locale = locale(encoding = "latin1"),
       na = c("NA", "n/a")) %>% 
  rename("edu" = "education/community_building") %>% 
  rename("tech" = "tech/STEM") %>% 
  rename("arts" = "arts/storytelling") %>% 
  filter(include == 1)
  
# as.Date requires a day so i added on bc yearmon sucks
# data$start_date <- as.yearmon(data$start_date,format="%m/%Y")
# data$end_date <- as.yearmon(data$end_date,format="%m/%Y")

data$start_date <- as.Date(data$start_date,format="%m/%d/%Y")
data$end_date <- as.Date(data$end_date,format="%m/%d/%Y")
```

```{r}
# for each month between Sep 2015 - Mar 2024
# get all rows where this month falls in between start and end (inclusive)
# sum all the rows in cols edu, tech, arts and divide by total 

#                     month      edu  tech  arts
# result should be   Jun 2022   0.33  0.66   0
```

```{r}
range_mo <- seq(as.Date("2015-09-01"), as.Date("2024-04-01"), "months") 
# print(range_months)

range_qu <- seq(as.Date("2015-09-01"), as.Date("2024-04-01"), "quarter") 
print(range_qu)

range_yr <- seq(as.Date("2015-01-01"), as.Date("2025-04-01"), "years") 
print(range_yr)
```

```{r}
# basically for each month, get all the rows whose time frame overlaps, and add up their column values, then divide by the total 

count_rows <- function(date) {
  # this only works for the month one
  temp <- data %>% filter(date >= data$start_date & date <= data$end_date) 
  # maybe i need all rows where either start or end date falls within curr and next date (for years). i could make 2 ranges that have an offset, and use both parameters
  total <- sum(rowSums(temp %>% select(c('edu', 'tech', 'arts'))))
  edu <- sum(temp$edu) / total
  tech <- sum(temp$tech) / total
  arts <- sum(temp$arts) / total
  return (c(edu, tech, arts))
}
```


```{r}
percentages_mo <- sapply(range_mo, count_rows)

percentages_qu <- sapply(range_qu, count_rows)
percentages_qu

percentages_yr <- sapply(range_yr, count_rows)
percentages_yr
```

I might want to give weights in the spreadsheet, like levels of time commitment, or manually weighting if something was tech and education but way more education. 
Could also add project pac.

```{r}
percentages <- percentages_qu
range <- range_qu
stackedAreaData <- data.frame(year=range,education=percentages[1,],tech=percentages[2,],arts=percentages[3,])

# we need this in long format for stack area chart in R ggplot2

data_long <- pivot_longer(stackedAreaData, cols = c(education, tech, arts), names_to = "category", values_to = "proportion")

print(data_long)
```



```{r}
# stacked area chart
stackedAreaQu <- ggplot(data_long, aes(x=year, y=proportion, fill=category)) + 
    geom_area(alpha=0.6 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) +
    theme_ipsum() + 
    ggtitle("My experience over time")
stackedAreaQu
```

```{r}
ggsave("stackedAreaQuarters.png")
```

I could also make a distinction between SWE and data tech
could add back in community

Would love to animate it so that it draws left to right
and prettify it so that it looks more flashy/modern