---
title: "Personal Experience Data Processing"
---

```{r}
library(readr)
library(zoo)
library(tidyr)
library(dplyr)
library(ggplot2)
library(viridis)
library(hrbrthemes)
```


```{r}
data <- read_csv("personalExperienceData.csv",
      locale = locale(encoding = "latin1"),
       na = c("NA", "n/a")) %>% 
  filter(include == 1)
  
# as.Date requires a day so i added on bc yearmon sucks
# data$start_date <- as.yearmon(data$start_date,format="%m/%Y")
# data$end_date <- as.yearmon(data$end_date,format="%m/%Y")

data$start_date <- as.Date(data$start_date,format="%m/%d/%Y")
data$end_date <- as.Date(data$end_date,format="%m/%d/%Y")
```
multiply the weights by the time commitment
```{r}
data$edu_w <- data$edu * data$time_level
data$arts_w <- data$arts * data$time_level
data$tech_w <- data$tech * data$time_level
```


for each [month / quarter / year] between Sep 2015 - Mar 2024
get all rows where this [month / quarter / year] falls in between start and end date (inclusive)
sum all the rows in cols edu, tech, arts (which are weighted) and divide by total

                   month      edu  tech  arts
result should be   Jun 2022   0.33  0.66   0

for each supplied date, get all the rows whose date range overlaps, and add up their column values for category weights, then divide by the total 
```{r}

count_rows <- function(date) {
  # this only works for the month one
  temp <- data %>% filter(date >= data$start_date & date <= data$end_date) 
  # maybe i need all rows where either start or end date falls within curr and next date (for years). i could make 2 ranges that have an offset, and use both parameters
  total <- sum(rowSums(temp %>% select(c('edu_w', 'tech_w', 'arts_w'))))
  edu <- sum(temp$edu_w) / total
  tech <- sum(temp$tech_w) / total
  arts <- sum(temp$arts_w) / total
  return (c(edu, tech, arts))
}
```

creating set of dates
```{r}
range_mo <- seq(as.Date("2015-09-01"), as.Date("2024-06-01"), "months") 

range_qu <- seq(as.Date("2015-09-01"), as.Date("2024-06-01"), "quarter") 

range_yr <- seq(as.Date("2015-01-01"), as.Date("2025-06-01"), "years") 
```

calculating percentages in each category with different time unit breakdowns, to see which makes most sense
```{r}
percentages_mo <- sapply(range_mo, count_rows)

percentages_qu <- sapply(range_qu, count_rows)

percentages_yr <- sapply(range_yr, count_rows)
```


Creating a long format dataset using the preferred time breakdown (quarters) to be used for stacked area chart in ggplot2
```{r}
# set time breakdown
percentages <- percentages_qu
range <- range_qu

stackedAreaData <- data.frame(year=range,education=percentages[1,],tech=percentages[2,],arts=percentages[3,])

# convert to long format
data_long <- pivot_longer(stackedAreaData, cols = c(education, tech, arts), names_to = "category", values_to = "proportion")

print(data_long)
```

```{r}
# custom palette
palette <- c("#C3AFFC", "#CDE8B5", "#FCBF4E") #FFC864
scales::show_col(palette)

text_color <- "#454545"
```


```{r}
# stacked area chart
stackedAreaQu <- ggplot(data_long, aes(x=year, y=proportion, fill=category)) + 
    geom_area(alpha=0.6 , size=.5, colour="white", show.legend = FALSE) +
    scale_x_date(expand = c(0, 0)) +
    scale_y_continuous(limits = c(0,1), expand = c(0, 0)) +
    scale_fill_manual(values = unname(palette)) +
    theme(text = element_text(size=12,  family="Avenir"), 
          plot.margin=unit(c(5,18,0,0), "mm"),
          aspect.ratio=0.7,
          panel.grid=element_blank(), 
          panel.background=element_rect(fill="white"),
          plot.background = element_rect(fill='transparent', color=NA),
          legend.background = element_rect(fill='transparent'), #transparent legend bg
          legend.box.background = element_rect(fill='transparent', linewidth = 0), #transparent legend panel
          axis.title.y = element_text(margin=margin(r = 15), color=text_color),
          axis.title.x = element_text(margin=margin(t = 10), color=text_color),
          axis.text = element_text(color=text_color),
          ) 
    # ggtitle("My experience over time")
stackedAreaQu
```

```{r}
ggsave("stackedAreaQuartersWeightedExt.png")
```

Future Ideas: 
- could also make a distinction between SWE and data tech
could add back in community
- would love to animate it so that it draws left to right
and prettify it so that it looks more flashy/modern